---
title: "p8105_hw2_hs3163"
author: "Hao Sun"
date: "9/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)

```

##Problem 1

This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website.

Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data
round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)


__Answer__
```{r Problem 1 part 1}
##Read Mr. Trash Wheel sheet, Ommiting the Column O and the header image.=
Mr_Trash_Wheel <- read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "Mr. Trash Wheel",range = "A2:N336")
Mr_Trash_Wheel<-Mr_Trash_Wheel%>%
##Remove the rows that don't contains row specific data
drop_na(Dumpster) %>%
##Change month into factor
  janitor::clean_names()%>%
  mutate(sports_balls = as.integer(sports_balls))
```

Read and clean precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}

Precipitation_2018<- read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2018 Precipitation",range = "A2:B7")
Precipitation_2017<- read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2017 Precipitation",range = "A2:B14")

```








