---
title: "p8105_hw2_hs3163"
author: "Hao Sun"
date: "9/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)

```

## Problem 1

This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website.

Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
use reasonable variable names
omit rows that do not include dumpster-specific data
round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)


__Answer__
```{r Problem 1 part 1}

##Read Mr. Trash Wheel sheet, Ommiting the Column O and the header image.=
Mr_Trash_Wheel <- read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "Mr. Trash Wheel",range = cell_cols("A:N"))
Mr_Trash_Wheel_cleaned <- Mr_Trash_Wheel%>%
##Remove the rows that don't contains row specific data
drop_na(Dumpster) %>%
##Change month into factor
  janitor::clean_names()%>%
  mutate(sports_balls = as.integer(sports_balls))%>%
##all the typr of trash are actually one variable
  pivot_longer(plastic_bottles:sports_balls,names_to = "trash_type", values_to = "trash_count")%>%
##Trash type shall be factor for easier data analysis
    mutate(trash_type = factor(trash_type)) %>%
###To avoid confusion, rename
  rename( all_type_weight_in_tons = weight_tons, 
          all_type_volumn_in_cubic_yards = volume_cubic_yards,
          all_type_homes_powered = homes_powered)%>%View

```

Read and clean precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
__Answer__
```{r Problem 1 part 2}

Precipitation_2018<- read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2018 Precipitation",range = "A2:B9")

Precipitation_2017<- read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2017 Precipitation",range = "A2:B14")
ncol(Precipitation_2018)

###Clean up the data for 2018
Precipitation_2018<-Precipitation_2018 %>% 
  rename( "Total Precipitation" = Total)%>%
  mutate(year = 2018)%>%
  mutate(Month = month.name[Month])%>%
  janitor::clean_names()

###Clean up the data for 2017
Precipitation_2017<-Precipitation_2017 %>% 
  rename( "Total Precipitation" = Total)%>%
  mutate(year = 2017)%>%
  mutate(Month = month.name[Month])%>%
  janitor::clean_names()

##Stack the two datas.
Precipitation_2017_2018 <-bind_rows(Precipitation_2017,Precipitation_2018)
Precipitation_2017_2018



```


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

__Answer__
The number of observations in the first resulting datasets is `r nrow(Mr_Trash_Wheel)`. The number of observations in the second resulting datasets is `r nrow(Precipitation_2017_2018)`. One key variable for the first table is the date, as the trash count for all the trash type except for glass bottles and cigarette increase till the middle of 2016 and then gradually drop. That said, another key variable will, therefore, be the type of trash as different types of trash have a different pattern on their trash count over time. 

The key variables to the second result table are the variable month and year, for as time progress the total precipitation changes along with it.

For key variables in terms of indexing, the key variable for both data set are year and month for they combined can be used to identified a unique record in the second table to into the first one.

The total precipitation in 2018 is `r filter(Precipitation_2017_2018,year == 2018)%>%pull(total_precipitation)%>%sum`. The median number of sports balls in a dumpster in 2017 is `r filter(Mr_Trash_Wheel_cleaned,year == 2017 & trash_type == "sports_balls")%>%pull(trash_count)%>%median`


##  Problem 2
First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

__Answer__

```{r Problem 2 part 1}
##Read the data set
pols_month<-read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv")%>%
##separate the mon variable.
separate(col = mon, into = c("year", "month","day"), sep = "-")%>%
##Change the month in to number, and then map the correspond month name on it.
  mutate(month = factor(month.name[as.numeric(month)]))%>%
##Since president is either gop or dem, in prez_gop there is misentered value :2, so prez_dem will be used as ref
  mutate(presidents = factor(prez_dem, labels = c("gop","dem")))%>%
  ##Since year is a number, change year into numeric
  mutate(year = as.numeric(year))%>%
##Remove the unneeded columns.
  select(-prez_dem,-prez_gop,-day)

```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

__Answer__


```{r Problem 2 part 2}
##Read the data set
snp<-read_csv(file = "./data/fivethirtyeight_datasets/snp.csv")%>%
##separate the date variable.
separate(col = date, into = c("month","day","year"), sep = "/")%>%
##Change the month in to number, and then map the correspond month name on it.
  mutate(month = factor(month.name[as.numeric(month)]))%>%
  ##Since year is a number, change year into numeric
  mutate(year = as.numeric(year))%>%
##Remove the unneeded columns and arrange the column.
  select(year, month, close)
```
Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

__Answer__

```{r Problem 2 part 3}
unemployment <-read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>%
pivot_longer(Jan:Dec,names_to = "month", values_to = "unemployment_rate")%>%
  ##Change the month into number
  mutate(month = match(month,month.abb)) %>%
  ##Change the month into full name
  mutate(month = factor(month.name[month])) %>%
  ##Change the year type into character
  rename(year = Year) %>%
  ##Rearrange the order
  select(year,month,unemployment_rate)

```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r Problem 2 part 4}
pols_snp_month<- left_join(pols_month,snp, by = c("year","month") )
pols_snp_unemployment_per_month <- left_join(pols_snp_month,unemployment,by = c("year","month")  )
pols_snp_unemployment_per_month
```
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).



__Answer__

The pols-month data set contains 822 observations of the number of democratic or republican politicians at president, governer, and senator levels for any given montns within the range of Jan 1947 to Jun 2015. Key variables are the year and months as the observations are "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem","rep_dem", and "presidents".

The pols-month data set contains 822 observations of the number of democratic or republican politicians at president, governer, and senator levels for any given montns within the range of Jan 1947 to Jun 2015. Key variables are the year and months as the observations are "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem","rep_dem", and "presidents".






##  Problem 3

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, the names of a categorical predictor and the case structure of string variables changed over time; you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will need to be removed (hint: google something like “dplyr remove duplicate rows” to get started).

```{r Problem 3 part 1}
Popular_Baby_Names <-read_csv(file = "./data/Popular_Baby_Names.csv") 

Popular_Baby_Names

Popular_Baby_Names<-Popular_Baby_Names%>%
##Cleanup the names
  janitor::clean_names()%>%
##First handle the difference in cases 
mutate(childs_first_name = str_to_sentence(childs_first_name))

##clean up the different format in ethincity
##Fix the ethnicity one by one
Popular_Baby_Names_Asians<-Popular_Baby_Names%>%  
  filter(str_detect(ethnicity,"ASIAN"))%>%
  mutate(ethnicity = "ASIAN AND PACIFIC ISLANDER")

Popular_Baby_Names_White<-Popular_Baby_Names%>%  
  filter(str_detect(ethnicity,"WHITE"))%>%
  mutate(ethnicity = "WHITE NON HISPANIC")


Popular_Baby_Names_Black<-Popular_Baby_Names%>%  
  filter(str_detect(ethnicity,"BLACK"))%>%
  mutate(ethnicity = "BLACK NON HISPANIC")


Popular_Baby_Names_Hisp<-Popular_Baby_Names%>%  
  filter(ethnicity == "HISPANIC")

##Bind the cleaned pieces together
Popular_Baby_Names_Cleaned<-bind_rows(Popular_Baby_Names_Asians,Popular_Baby_Names_White,Popular_Baby_Names_Black,Popular_Baby_Names_Hisp)%>%
##Remove duplicated rows
distinct()%>%
##Change the approperate variables into factors
mutate(ethnicity = factor(ethnicity)) %>%
  mutate(gender = factor(gender))
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.



```{r Problem 3 part 2}
###Finding the popularity of Olivia
##Filter out the portion with only olivia
Popularity_of_Olivia<-Popular_Baby_Names_Cleaned%>%
  filter(childs_first_name == "Olivia")%>%
    filter(gender == "FEMALE")%>%
##Select only the needed column
  select(ethnicity,year_of_birth,rank)%>%
##Make the table wider
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank")%>%
  knitr::kable(caption = "Table 1.The popularity of Olivia")
Popularity_of_Olivia

```


```{r Problem 3 part 2 Cont}
####Finding the most popular boy's name
Most_popular_boy_s_name<-Popular_Baby_Names_Cleaned%>%
  filter(rank == "1")%>%
  filter(gender == "MALE")%>%
##Select only the needed column
  select(ethnicity,year_of_birth,childs_first_name)%>%
##Make the table wider
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "childs_first_name")%>%
knitr::kable(caption = "Table 2.The most popular boy's name")
Most_popular_boy_s_name
```

__Answer__
The most popular boys name is for Asian and Pacific Islander as well as Hispanic is Jayden, for White is Joseph, for Black are Jayden, Ethan, and Noah. Therefore, the most popular boy's name over time is Jayden. 

_______________________________________________________________________________________________________________



Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).


```{r Problem 3 part 3}
##Filter out with only male, white non-hispanic children born in 2016

Popular_Baby_Names_Cleaned%>%
  filter( year_of_birth  == 2016 )%>%
    filter(gender == "MALE") %>%
    filter(ethnicity == "WHITE NON HISPANIC")%>%
## Make the plot
  ggplot(aes(x = rank, y = count)) + 
  geom_point() +
ggtitle("For 2016 White boy, the popularity of a name against the number of children with it")+
labs(y="Number of children with a name", x = "Rank in popularity of a name")

```

